Preliminary Insights into Indigenous Subsistence and Colonial Encounters: A Zooarchaeological Perspective from Kiwulan in Northeastern Taiwan

### 2025-06-15 analysis/paper/paper.qmd
editing

Surface modifications, including cut marks, burning damage, and fractures, were identified on deer bones from CL6 to CL2, with variations in their proportions (@tbl-deer-modify). Cut marks appeared on 54% of bones in CL2, around 30% in earlier periods (CL6 to CL4), and at a lower proportion of 23% in CL3. Most cut marks were located on major limb joints and long bones, indicating butchering activities related to disarticulation and meat removal. At least 50% of joints in CL4 presents cut marks as shown in @fig-deer-cutmark. A similar pattern was observed in CL5, but with a lower proportion of cut marks on the ankle, knee, and hip compared to CL4. Direct comparisons across all layers are challenging since not all joints were found in CL6, CL3, and CL2. Nevertheless, the overall high frequency of joints with cut marks suggests intensive deer processing throughout the site's occupation. In addition, cut marks on the lower limbs, particularly shallow, elongated traces on the medial and lateral sides of bones, are likely associated with skinning (@fig-cutmark-img: e). These marks were primarily observed on metapodials in CL4 (n = 6), CL5 (n = 4), CL6 (n = 1), and CL3 (n = 1), with additional occurrences on first phalanges in CL4 (n = 1) and CL5 (n = 1). Some joints displayed clean, smooth cut surfaces from chopping (@fig-cutmark-img: a, c), accompanied by narrow V-shaped cut marks observed on most bones, indicating the use of metal tools [@Greenfield1999].

>>>>

Surface modifications, including cut marks, burning damage, and fractures, were identified on deer bones from CL6 to CL2, with variations in their proportions (@tbl-deer-modify). Cut marks appeared on 54% of bones in CL2, around 30% in earlier periods (CL6 to CL4), and at a lower proportion of 23% in CL3. Most cut marks were located on major limb joints and long bones, indicating butchering activities related to disarticulation and meat removal. At least 50% of joints in CL4 presents cut marks as shown in @fig-deer-cutmark. A similar pattern was observed in CL5, but with a lower proportion of cut marks on the ankle, knee, and hip compared to CL4. Direct comparisons across all layers are challenging since not all joints were found in CL6, CL3, and CL2. Nevertheless, the overall high frequency of joints with cut marks suggests intensive deer processing throughout the site's occupation. In addition, cut marks on the lower limbs, particularly shallow, elongated traces on the medial and lateral sides of metapodials (@fig-cutmark-img: e), are consistent with marks left by skinning based on documented butchery experiments [@Costamagno2019; @Val2011]. These skinning marks were identified on metapodials from CL4 (n = 6), CL5 (n = 4), CL6 (n = 1), and CL3 (n = 1), with additional occurrences on first phalanges from CL4 (n = 1) and CL5 (n = 1). Some joints displayed clean, smooth cut surfaces from chopping (@fig-cutmark-img: a, c), accompanied by narrow V-shaped cut marks observed on most bones, indicating the use of metal tools [@Greenfield1999].
